---
title: "RDM homework - Analysis report"
author: "Helene Langet"
format: html
editor: visual
execute:
  echo: false
  warning: false
---

```{r}
# Load required libraries

# Load the tidyverse collection of packages for data manipulation and visualization
library(tidyverse)

# Load the haven package to read and work with Stata, SPSS, and SAS data files
library(haven)

# Load the gtsummary package to create publication-ready summary tables
# Useful for descriptive statistics and model result tables with clean formatting
library(gtsummary)
```

```{r}
source("1_cleaning_script.R")
```

```{r}
# Load the (clean) household-level data 
#household_df <- haven::read_dta("../data/cleanR_household_with_village.dta")

# Load the (clean) household-level data 
# ...
```

# About the National Malaria Indicator Survey 2019/2020

Cross-sectional household survey in all 22 provinces of Papua New Guinea.

-   5 randomly sampled villages/province
-   ca. 30 random households/village
-   Structured interviews
-   Finger-prick blood sample for malaria diagnosis by Rapid Diagnostic Test (RDT) and light microscopy

# Shell tables

## Shell 1

## Shell 2

```{r}
household_with_village_df |>
  dplyr::select(altitude) |>
  gtsummary::tbl_summary(label = list("altitude" ~ "Altitude (m)"))
```

## Shell 3

```{r}
clean_individual_df |>
  dplyr::select(age) |>
  gtsummary::tbl_summary()
```

# Question 1

How many household interviews were conducted per month?

```{r}
count_df <- household_with_village_df |>
  dplyr::group_by(month) |>
  dplyr::count()
count_df |> 
  knitr::kable()
```

Instructions:

-   Step 1: Add color to bars
-   Step 2: Add data labels to the bars
-   Step 3: Adjust the y-axis limits
-   Step 4: Improve x-axis date formatting
-   Step 5: Add labels to the axes
-   Step 6: Highlight year transitions with a vertical line
-   Step 7: Apply a minimal theme
-   Step 8: Customize theme elements for cleaner look
-   Step 9: Add a caption
-   Step 10: Adjust the dimensions of the figure so that its width is twice its height

```{r}
count_df |>
  ggplot(mapping = ggplot2::aes(x = month, y = n)) +
  ggplot2::geom_bar(stat = "identity", fill = "blue") +
  ggplot2::geom_text(mapping = ggplot2::aes(label = n), vjust = -0.5, color = "black") +
  ggplot2::ylim(0, M) +
  ggplot2::scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  ggplot2::labs(x = "Month",
                y = "Number of household interviews") +
  ggplot2::geom_vline(xintercept = as.numeric(as.Date(paste0(unique(format(household_merge_df$month, "%Y")), "-01-01"))),
                      linetype = "dashed") +
  ggplot2::theme_minimal() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
                 panel.grid.major.x = ggplot2::element_blank(),
                 panel.grid.minor.x = ggplot2::element_blank(),
                 panel.grid.major.y = ggplot2::element_blank(),
                 panel.grid.minor.y = ggplot2::element_blank(),
                 axis.text.y = ggplot2::element_blank(),
                 axis.ticks.y = ggplot2::element_blank())
```

# Question 2

Are members of households that own at least one long-lasting insecticidal net (LLIN) less likely to be infected with malaria parasites?

```{r}
cross_tab <- table(individual_merge_df$mosquitonet, individual_merge_df$microscopy_res)

# chi-square test
chi_sq_test <- prop.test(cross_tab, correct = FALSE)
chi_sq_test
```

# Question 3

Does prevalence of Plasmodium vivax and P. falciparum differ between altitudinal zones?

# Question 4

What is the sensitivity and specificity of RDTs compared to microscopy (any species)?
